\section{Current Work and Living Status}

This document is intended as a living Overleaf document for the optimized and explained CAGE Challenge 4 fork. It is not a replacement for the original competition documentation; it is a research-oriented technical companion that records what was changed, what was fixed, how the current agents work, and which open questions remain.

\subsection{Active Work Items}

\begin{longtable}{p{0.26\textwidth}p{0.14\textwidth}p{0.50\textwidth}}
\toprule
\textbf{Item} & \textbf{Status} & \textbf{Notes} \\
\midrule
Validate current best heuristic version & Open & The repository contains v11a and v11b submission adapters. v11a is documented as the current strongest measured heuristic in the architecture notes; v11b exists as a coordinated messaging variant and should be benchmarked before being described as the best version. \\
Refresh benchmark table & Open & README, architecture notes, and research-gap notes use different baselines and evaluation sample sizes. Consolidate around one benchmark protocol: seed, episodes, episode length, reward aggregation, and submission path. \\
Document optimized environment fixes & In progress & Speed optimizations and simulation-audit findings are documented locally. The exact set of fixes that are already implemented should be cross-checked against the code and commit history. \\
Explain MARL training pipeline & Drafted & This document summarizes the graph-based PPO/GNN approach from \texttt{marl/README.md}, \texttt{marl/models/cage4.py}, and \texttt{marl/wrappers/graph\_wrapper.py}. \\
Unify submission selection & Implemented & \texttt{explain.py} can now load registered agents with \texttt{--agent NAME} from \texttt{submissions/registry.json}. Current entries include \texttt{heuristic\_v11a}, \texttt{heuristic\_v11b}, \texttt{marl\_test\_cpu}, and \texttt{gpu\_test}. \\
Make SHAP runs scalable & Implemented & Large explainability runs can evaluate all episodes while sampling SHAP rows through \texttt{--shap-episode-stride}, \texttt{--shap-step-stride}, and \texttt{--shap-max-rows-per-profile}. A 100 episode by 500 step heuristic run completed with 50,000 logged steps and a capped SHAP phase. \\
Overleaf polish & Open & Add figures, final benchmark plots, glossary, and stable citations once the text is reviewed. \\
\bottomrule
\end{longtable}

\subsection{Document Conventions}

\begin{itemize}
  \item ``Original CC4'' refers to the public CAGE Challenge 4 release and documentation.
  \item ``Optimized fork'' refers to this repository.
  \item ``Heuristic agent'' refers to the rule-based EnterpriseHeuristicAgent family.
  \item ``MARL agent'' refers to the graph-based PPO agent under \texttt{marl/}.
  \item Registered submissions should be referenced by their \texttt{--agent} name when possible.
  \item Performance claims should always include seed, number of episodes, and submission path.
\end{itemize}
