\section{Optimized Fork: Purpose and Changes}

The optimized fork keeps the original CC4 task structure but adds performance improvements, simulation fixes, analysis tooling, heuristic submissions, and a graph-based MARL pipeline. The repository README describes the fork as an optimized research environment for autonomous cyber-defence, with faster training, simulation bug fixes, and strong heuristic baselines.

\subsection{High-Level Additions}

\begin{itemize}
  \item \textbf{Optimized simulation path}: targeted changes reduce reset and step overhead.
  \item \textbf{Bug and consistency audits}: local reports identify issues in phishing, FSM fallback behavior, restore semantics, detection events, and observability.
  \item \textbf{Heuristic blue agents}: the EnterpriseHeuristicAgent family encodes a strong rule-based defence strategy.
  \item \textbf{MARL pipeline}: the \texttt{marl/} directory implements graph observations, GNN actor-critic networks, PPO training, evaluation, and explainability hooks.
  \item \textbf{Explainability tooling}: reward decomposition, action logging, SHAP-oriented features, and plotting scripts are wired into evaluation runs.
\end{itemize}

\subsection{Implemented Reward Variants}

The fork adds reward-mode switches through environment variables read by
\texttt{BlueRewardMachine}. The active mode is selected with
\texttt{CYBORG\_PHASE\_REWARD\_MODE}; blue action shaping is separately enabled
with \texttt{CYBORG\_REWARD\_BLUE=1}. This matters for comparable experiments:
results from different modes should not be mixed in the same benchmark table.

\begin{longtable}{p{0.22\textwidth}p{0.68\textwidth}}
\toprule
\textbf{Mode} & \textbf{Semantics} \\
\midrule
\texttt{default} & Uses the original LWF, ASF, and RIA phase table. Green failures contribute when \texttt{GreenLocalWork} or \texttt{GreenAccessService} fails. Red contributes only on successful \texttt{Impact}. \\
\texttt{contractor\_off} & Copies \texttt{default}, then sets \texttt{contractor\_network\_subnet} to zero for LWF, ASF, and RIA in every mission phase. This removes contractor-zone reward pressure. \\
\texttt{red\_only} & Sets LWF and ASF to zero and keeps red penalties. It also decomposes successful red progress into \texttt{EXP}, \texttt{PRIV}, and \texttt{DEG} terms for exploit, privilege escalation, and service degradation. \\
\bottomrule
\end{longtable}

In \texttt{red\_only}, the extra red terms are derived directly from each
subnet's RIA value:

\begin{center}
\begin{tabular}{lll}
\toprule
\textbf{Term} & \textbf{Triggered by} & \textbf{Value} \\
\midrule
\texttt{EXP} & successful \texttt{ExploitRemoteService} & $0.7 \times \texttt{RIA}$ \\
\texttt{PRIV} & successful \texttt{PrivilegeEscalate} & $0.8 \times \texttt{RIA}$ \\
\texttt{DEG} & successful \texttt{DegradeServices} & $0.6 \times \texttt{RIA}$ \\
\bottomrule
\end{tabular}
\end{center}

When \texttt{CYBORG\_REWARD\_BLUE=1}, successful blue \texttt{Analyse} actions
receive a small shaping reward of $+0.1$. In the default repository state this
flag is off, so blue actions do not normally add positive reward.

\subsection{Reward Logging}

The fork logs reward decomposition to JSON lines from the reward machine. The
log path defaults to \texttt{reward\_log.jsonl} and can be changed with
\texttt{CYBORG\_REWARD\_LOG\_PATH}. Optional metadata fields are read from
\texttt{CYBORG\_ATTACK\_PROFILE} and \texttt{CYBORG\_RUN\_TAG}. Each entry
contains episode, step, mission phase, per-subnet reward components, and total
reward. This is the preferred source for reward-decomposition plots and for
checking whether a benchmark was dominated by green failures, exploit progress,
privilege escalation, degradation, or final impact.

\subsection{Performance Optimizations}

The local speed report records three waves of optimization. The measured result for the documented run was 87.7 steps/second, compared with 64.7 steps/second before the optimization pass. The report attributes the improvement to changes such as:

\begin{itemize}
  \item pre-allocated observation buffers in \texttt{BlueFlatWrapper};
  \item wireless-topology and connected-agent caches in \texttt{SimulationController};
  \item static constants in \texttt{EnterpriseScenarioGenerator};
  \item replacing list membership with set/dictionary based lookups for PID and session indexing;
  \item faster clone paths for process and network-connection objects;
  \item scenario-pool wiring through the public \texttt{CybORG} constructor.
\end{itemize}

The optimization goal is behavior preservation. Whenever a performance change is made, benchmark reward should remain stable under the same seed and episode protocol.

\subsection{Simulation Issues Identified}

The simulation audit identifies several issues worth tracking. Some are bugs, some are questionable modelling choices, and some are deliberate simplifications that should be documented.

\begin{longtable}{p{0.18\textwidth}p{0.36\textwidth}p{0.36\textwidth}}
\toprule
\textbf{ID} & \textbf{Issue} & \textbf{Impact} \\
\midrule
C1 & \texttt{PhishingEmail} uses physical routability rather than the firewall blocking check. & Firewall rules do not affect phishing delivery. This may be realistic as an email abstraction, but the implementation path is not explicit. \\
C2 & \texttt{PhishingEmail} fallback loop can fail to shrink the candidate list. & Potential non-termination or unintended cross-zone source selection. \\
C3 & FSM red-agent fallback references a missing method when known hosts are exhausted. & Potential runtime crash in certain dead-end states. \\
B1 & Historical notes said \texttt{Remove} stopped suspicious processes but did not remove files. & In the current fork implementation, \texttt{Remove} also removes high-density unsigned malware files. Restore is still required to fully evict privileged/root red sessions. \\
B2 & Restore can clear pending detection events. & Blue can miss signals on Restore steps. \\
B3 & Multi-hop red sessions can become orphaned after Restore. & Potential inconsistent red session state. \\
\bottomrule
\end{longtable}

\subsection{Open Implementation Check}

This document currently records findings from the local reports. Before publication, each issue should be marked as:
\begin{itemize}
  \item \textbf{fixed}: code was changed and regression-tested;
  \item \textbf{documented design choice}: behavior remains but is explained;
  \item \textbf{open}: known issue still exists.
\end{itemize}
