\section{Environment Model}

\subsection{Network and Agents}

CC4 uses five blue agents. Four agents defend a single restricted or operational zone; the fifth protects the public-access, admin, and office areas. A simplified assignment is:

\begin{center}
\begin{tabular}{lll}
\toprule
\textbf{Agent} & \textbf{Controlled area} & \textbf{Role} \\
\midrule
\texttt{blue\_agent\_0} & Restricted Zone A & guard upstream path to OZA \\
\texttt{blue\_agent\_1} & Operational Zone A & protect phase-1 mission target \\
\texttt{blue\_agent\_2} & Restricted Zone B & guard upstream path to OZB \\
\texttt{blue\_agent\_3} & Operational Zone B & protect phase-2 mission target \\
\texttt{blue\_agent\_4} & Public access, admin, office & guard HQ-facing subnets \\
\bottomrule
\end{tabular}
\end{center}

\subsection{Generated Topology}

The scenario generator creates eight enterprise subnets plus the internet
subnet. Each non-internet subnet contains one router, between three and ten
user hosts, and between one and six server hosts. The internet subnet contains
only \texttt{root\_internet\_host\_0}. Host operating systems are Linux variants
in the current generator. Each non-router host has SSH; operational-zone hosts
also have \texttt{OTService}; optional services are Apache, MySQL, and SMTP.

\begin{longtable}{p{0.30\textwidth}p{0.56\textwidth}}
\toprule
\textbf{Subnet} & \textbf{Primary links in generated topology} \\
\midrule
\texttt{internet\_subnet} & connected to restricted A, restricted B, contractor, and public-access routers \\
\texttt{restricted\_zone\_a\_subnet} & internet and operational zone A \\
\texttt{operational\_zone\_a\_subnet} & restricted zone A only \\
\texttt{restricted\_zone\_b\_subnet} & internet and operational zone B \\
\texttt{operational\_zone\_b\_subnet} & restricted zone B only \\
\texttt{contractor\_network\_subnet} & internet-facing, with server-level links to restricted A, restricted B, and public access \\
\texttt{public\_access\_zone\_subnet} & internet, admin network, office network, and contractor-facing paths \\
\texttt{admin\_network\_subnet} & public-access network and office network \\
\texttt{office\_network\_subnet} & public-access network and admin network \\
\bottomrule
\end{longtable}

\subsection{Agent Action Permissions}

Action permissions are generated centrally in
\texttt{EnterpriseScenarioGenerator}. Blue agents receive the same action
classes, but only over their assigned subnets. Red agents receive the same red
action classes, but only one red agent starts active: the red agent initialized
in the contractor network. Green agents are created for each generated user
host.

\begin{longtable}{p{0.16\textwidth}p{0.26\textwidth}p{0.48\textwidth}}
\toprule
\textbf{Team} & \textbf{Agents} & \textbf{Action classes} \\
\midrule
Blue & \texttt{blue\_agent\_0}--\texttt{blue\_agent\_4} & \texttt{AllowTrafficZone}, \texttt{BlockTrafficZone}, \texttt{Monitor}, \texttt{Analyse}, \texttt{Restore}, \texttt{Remove}, \texttt{DeployDecoy}, \texttt{Sleep} \\
Red & \texttt{red\_agent\_0}--\texttt{red\_agent\_5} & \texttt{DiscoverRemoteSystems}, \texttt{AggressiveServiceDiscovery}, \texttt{StealthServiceDiscovery}, \texttt{ExploitRemoteService}, \texttt{PrivilegeEscalate}, \texttt{DegradeServices}, \texttt{DiscoverDeception}, \texttt{Impact}, \texttt{Withdraw}, \texttt{Sleep} \\
Green & one green agent per user host & \texttt{GreenAccessService}, \texttt{GreenLocalWork}, \texttt{Sleep}; when generated with \texttt{SleepAgent}, only \texttt{Sleep} \\
\bottomrule
\end{longtable}

\subsection{Action Costs and Durations}

The base action duration is one step and the base action cost is zero. Only a
small number of actions override these defaults. The table below lists the
high-level CC4 actions used by the generated scenario, not every low-level
helper action hidden inside them.

\begin{longtable}{p{0.22\textwidth}p{0.11\textwidth}p{0.11\textwidth}p{0.46\textwidth}}
\toprule
\textbf{Action} & \textbf{Team} & \textbf{Duration} & \textbf{Direct cost and effect} \\
\midrule
\texttt{Monitor} & Blue & 1 & cost 0; updates observation from local Velociraptor view. \\
\texttt{Analyse} & Blue & 2 & cost 0 unless blue shaping is enabled; checks a host with density/signature artefacts. \\
\texttt{Remove} & Blue & 3 & cost 0; kills suspicious non-privileged red processes and, in this fork, removes high-density unsigned malware files. Root sessions survive. \\
\texttt{Restore} & Blue & 5 & cost -1; reimages the host through \texttt{RestoreFromBackup} and is the reliable eviction path for privileged sessions. \\
\texttt{DeployDecoy} & Blue & 2 & cost 0; deploys a compatible exploit decoy process such as Apache, Tomcat, Haraka, or Vsftpd. \\
\texttt{BlockTrafficZone} & Blue & 1 & cost 0 direct; adds a subnet block. Indirect reward impact comes from green access failures and blocked red paths. \\
\texttt{AllowTrafficZone} & Blue & 1 & cost 0; removes a matching subnet block. \\
\texttt{DiscoverRemoteSystems} & Red & 1 & cost 0; pingsweeps a target subnet for active host IPs. \\
\texttt{AggressiveServiceDiscovery} & Red & 1 & cost 0; port scan with detection rate 0.75. \\
\texttt{StealthServiceDiscovery} & Red & 3 & cost 0; port scan with detection rate 0.25. \\
\texttt{ExploitRemoteService} & Red & 4 & cost 0; attempts a service exploit against known ports and creates a user-level red shell on success. \\
\texttt{PrivilegeEscalate} & Red & 2 & cost 0; upgrades a user shell to privileged/root access where possible. \\
\texttt{DegradeServices} & Red & 2 & cost 0; with privileged access, degrades all active services on the host. \\
\texttt{DiscoverDeception} & Red & 2 & cost 0; probes for decoy services with true-positive rate 0.5 and false-positive rate 0.1. \\
\texttt{Impact} & Red & 2 & cost 0; with privileged access, stops \texttt{OTService} and increments host impact count. \\
\texttt{Withdraw} & Red & 1 & cost 0; removes red sessions from a target host when routed. \\
\texttt{GreenAccessService} & Green & 1 & cost 0; tries to reach an allowed server service; failure can score ASF in non-\texttt{red\_only} modes. \\
\texttt{GreenLocalWork} & Green & 1 & cost 0; uses local services, may create false positives, and may trigger phishing via user error; failure can score LWF in non-\texttt{red\_only} modes. \\
\texttt{Sleep} & All & 1 & cost 0; no operational effect. \\
\texttt{InvalidAction} & Any & 1 & cost -0.1; returned by invalid action translation or execution paths. \\
\bottomrule
\end{longtable}

\subsection{Mission Phases}

Episodes run for 500 steps. The heuristic documentation uses the following phase boundaries:

\begin{itemize}
  \item Phase 0: steps 0--166, pre-planning/general operation.
  \item Phase 1: steps 167--333, Zone A becomes the high-priority mission area.
  \item Phase 2: steps 334--499, Zone B becomes the high-priority mission area.
\end{itemize}

The current phase is exposed in the observation. Agents must adapt both host response and traffic policy to the phase.

\subsection{Red Kill Chain}

The practical red chain observed by the defensive agents is:

\begin{lstlisting}
PhishingEmail -> Discover -> ServiceScan -> ExploitRemoteService
              -> PrivilegeEscalate -> DegradeServices / Impact
\end{lstlisting}

Two signatures are especially important in this fork:
\begin{itemize}
  \item \texttt{ExploitRemoteService} drops high-density unsigned command files.
  \item \texttt{PrivilegeEscalate} drops high-density unsigned escalation files.
\end{itemize}

These file artifacts motivate \texttt{BlueFlatWrapperV2}'s malicious-file bit.

\subsection{Observation Layout for the Heuristic Agent}

The heuristic agent consumes a flat vector from \texttt{BlueFlatWrapperV2}. Conceptually, the vector contains:

\begin{lstlisting}
[ phase | subnet blocks | subnet comms policy | process flags
  | connection flags | peer messages | malicious-file flags ]
\end{lstlisting}

The appended malicious-file flags are one bit per controlled host. A flag is set when a host contains a file with density at least 0.9 and no valid signature. Local notes state that green activity does not generate these files, making this a high-precision signal in the current environment.
