\section{Heuristic Agent}

\subsection{Current Implementations}

The repository contains submission adapters for \texttt{EnterpriseHeuristicAgentV11a} and \texttt{EnterpriseHeuristicAgentV11b}. The v11a adapter describes the agent as a rule-based priority heuristic with Restore-only threat response and preemptive Operational Zone blocking. The v11b adapter adds a redesigned messaging protocol for coordinated blocking.

Because local documents report mixed results, this draft treats v11a as the best documented baseline and v11b as an experimental messaging variant until a fresh benchmark confirms otherwise.

\subsection{Design Philosophy}

The heuristic is deterministic. At every step, it evaluates a priority cascade and executes the first valid action. There is no learned value function and no stochastic exploration. The strength of the agent comes from encoding environment-specific timing facts:

\begin{itemize}
  \item \textbf{Restore-biased response}: \texttt{Remove} now clears high-density unsigned malware files in this fork, but it still cannot evict privileged/root red sessions. \texttt{Restore} remains the reliable eviction path after privilege escalation.
  \item \textbf{Decoy saturation}: each host is filled with up to three decoys, increasing the chance that blind red exploitation wastes time.
  \item \textbf{Malicious-file detection}: high-density unsigned files are treated as strong evidence of compromise.
  \item \textbf{Flag-age filtering}: transient process flags are delayed to avoid green false positives.
  \item \textbf{Phase-aware blocking}: operational-zone paths are blocked around phase transitions to avoid windows where red can move while blue is busy.
\end{itemize}

\subsection{v11a: Preemptive Operational-Zone Blocking}

The v11a source states that the relevant transition windows are:

\begin{itemize}
  \item before Phase 1, block \texttt{restricted\_zone\_a\_subnet -> operational\_zone\_a\_subnet};
  \item before Phase 2, block \texttt{restricted\_zone\_b\_subnet -> operational\_zone\_b\_subnet}.
\end{itemize}

The rationale is timing. If the agent waits until a phase transition is visible, it may be occupied by higher-priority Restore actions. Preemptive blocking pays a small service-availability cost before the phase but avoids a high-risk open path at the beginning of the active mission phase.

\subsection{v11b: Messaging Redesign}

The v11b source describes a new 8-bit message format with fields for red detection, root-level suspicion, restore-busy status, zone-clear status, threat count, and block requests. Its goal is to let upstream agents ask downstream operational-zone agents to block traffic from compromised zones.

The research notes warn that earlier messaging designs showed little measurable value. The likely reasons are:
\begin{itemize}
  \item \texttt{comms\_policy} already provides much of the useful coordination;
  \item only a small subset of phase-agent combinations has useful upstream dependency;
  \item one-step message delay is large relative to red action timing.
\end{itemize}

\subsection{Submission Adapter}

The heuristic submission wraps the inner agent with a \texttt{BlueFlatWrapperV2}-based environment. The adapter bridges the official \texttt{evaluation.py} interface by:

\begin{itemize}
  \item retrieving action masks from the wrapper;
  \item forwarding the action index returned by the heuristic;
  \item storing outgoing messages from each agent;
  \item injecting those messages into the next environment step, because the official evaluator does not pass them directly.
\end{itemize}

\subsection{SHAP Feature Logging}

The v11b submission adapter now emits SHAP-oriented metadata through the wrapper \texttt{info} dictionary. For each blue agent step, the metadata includes the chosen action type and a semantic feature dictionary. This makes the heuristic explainability path structurally closer to the MARL path, which already emits \texttt{shap\_features} from the graph wrapper.

The heuristic SHAP pipeline trains a surrogate classifier on these recorded state-action rows. Categorical fields, such as previous action labels, are one-hot encoded before surrogate training. The output tables mirror the MARL SHAP output names:

\begin{itemize}
  \item \texttt{mean\_abs\_shap\_by\_feature.csv};
  \item \texttt{mean\_abs\_shap\_by\_feature\_group.csv};
  \item per-action-class feature and feature-group tables.
\end{itemize}

This does not make heuristic and MARL explanations mathematically identical, because they still use different policy representations and surrogate models. It does make the final CSV schema comparable enough to inspect feature and feature-group importance under the same red profile, seed, episode count, and SHAP sampling configuration.
