\section{Workflows}

\subsection{Comparable Evaluation Template}

For new experiments, prefer one parameterized command template over editing many
single-use scripts. The current repository already has this shape in
\texttt{explain.py}: it accepts seed, episode count, episode length, output
folder, submission path, reward mode, red-profile selection, and SHAP flags.
Registered agents should be selected with \texttt{--agent NAME} when possible,
because the registry captures the submission path, heuristic/MARL type, checkpoint
prefix, and checkpoint directory settings in one place.

\begin{lstlisting}
python explain.py \
  --agent heuristic_v11b \
  --mode single \
  --strategy single \
  --single_profile fsm_default \
  --max-eps 100 \
  --episode-length 500 \
  --seed 42 \
  --phase_reward_mode default \
  --output Results/EVAL_NAME
\end{lstlisting}

The fields that must be fixed before comparing two agents are:

\begin{itemize}
  \item \texttt{--seed}: controls scenario generation and stochastic choices.
  \item \texttt{--max-eps}: number of episodes.
  \item \texttt{--episode-length}: normally 500 for CC4.
  \item \texttt{--phase\_reward\_mode}: \texttt{default}, \texttt{contractor\_off}, or \texttt{red\_only}.
  \item \texttt{--single\_profile} or \texttt{--profile\_weights}: red attack distribution.
  \item \texttt{--agent}: exact registered agent implementation under test.
  \item \texttt{--output}: run-specific output directory; never reuse for another run.
\end{itemize}

The available registry entries can be inspected with:

\begin{lstlisting}
python explain.py --list-agents
\end{lstlisting}

For reward-decomposition experiments, set the output path and reward mode in the
same command and keep the generated JSONL reward log with the scalar summary.
For explainability experiments, add \texttt{--shap} plus fixed
\texttt{--shap-background-samples}, \texttt{--shap-explain-samples}, and
\texttt{--shap-random-state}.

\subsection{Heuristic Evaluation}

The repository README gives a direct evaluation pattern through the official
submission evaluator:

\begin{lstlisting}
python -c "
import sys; sys.path.insert(0, 'CybORG/Evaluation/submission')
from submission import Submission
from CybORG.Evaluation.evaluation import run_evaluation
run_evaluation(Submission, log_path='Results/v11a/', max_eps=100, seed=42)
"
\end{lstlisting}

For v11b, adjust the inserted submission path to \texttt{CybORG/Evaluation/submission\_v11b}.

For comparable fork experiments, the equivalent \texttt{explain.py} call is
usually preferable because it records action logs, reward decomposition, scalar
summaries, and optional SHAP artefacts in a single run directory:

\begin{lstlisting}
python explain.py \
  --agent heuristic_v11b \
  --mode single \
  --strategy single \
  --single_profile fsm_default \
  --max-eps 100 \
  --episode-length 500 \
  --seed 42 \
  --phase_reward_mode red_only \
  --output Results/v11b_red_only_fsm_default_seed42
\end{lstlisting}

\subsection{MARL Smoke Test}

The local workflow document recommends:

\begin{lstlisting}
source .venv/bin/activate
python marl/train.py mini_curriculum_smoke --mini \
  --strategy curriculum --warmup_updates 1 \
  --log_per_profile --device cpu
\end{lstlisting}

Expected outputs include reward logs under \texttt{logs/} and per-agent checkpoints under \texttt{checkpoints/}.

\subsection{Explainability Run}

The explainability pipeline can run a short single-profile pass:

\begin{lstlisting}
python explain.py \
  --agent marl_test_cpu \
  --mode single \
  --strategy single \
  --single_profile fsm_default \
  --max-eps 1 \
  --episode-length 5 \
  --shap \
  --seed 1337 \
  --phase_reward_mode red_only \
  --output SmokeResults/FULL_RUN
\end{lstlisting}

Expected outputs include reward logs, scalar reward summaries, action logs, and SHAP analysis folders.

\subsection{Large SHAP Sampling Runs}

Full CC4 explainability runs can be large: 100 episodes with 500 steps produce
50,000 environment steps for a single red profile. Computing model-agnostic SHAP
values on every recorded row is unnecessarily expensive. The current
\texttt{explain.py} implementation therefore separates evaluation size from SHAP
sample size:

\begin{itemize}
  \item \texttt{--shap-episode-stride}: keep SHAP rows only from every $N$th episode;
  \item \texttt{--shap-step-stride}: keep SHAP rows only from every $N$th step;
  \item \texttt{--shap-max-rows-per-profile}: deterministic cap before surrogate training and SHAP.
\end{itemize}

Example command for a full heuristic evaluation with bounded SHAP cost:

\begin{lstlisting}
python explain.py \
  --agent heuristic_v11b \
  --mode single \
  --single_profile fsm_default \
  --max-eps 100 \
  --episode-length 500 \
  --seed 1337 \
  --shap \
  --shap-step-stride 10 \
  --shap-max-rows-per-profile 2000 \
  --shap-background-samples 50 \
  --shap-explain-samples 100 \
  --output Results/shap_large_sampling
\end{lstlisting}

In a local validation run, this command evaluated all 100 episodes and logged
50,000 action/reward rows. The final SHAP phase used 2,000 sampled training rows
and 100 explained samples, completing in under one minute after the simulation
finished. The resulting SHAP folders contain both per-action plots and comparable
CSV summaries by feature and feature group.
